---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

\newpage

# Supplemental Materials for Data Rich but Model Resistant: An Evaluation of Data-Limited Methods to Manage Fisheries with Failed Age-based Stock Assessments

  
Christopher M. Legault^1^, John Wiedenmann^2^ Jonathan J. Deroba^1^, Gavin Fay^3^, Timothy J. Miller^1^, Elizabeth N. Brooks^1^, Richard J. Bell^4^, Joseph A. Langan^5^, Jamie M. Cournane^6^, Andrew W. Jones^1^, Brandon Muffley^7^


^1^National Marine Fisheries Service, Northeast Fisheries Science Center, Woods Hole, MA, USA  
^2^Department of Ecology, Evolution and Natural Resources, Rutgers, The State University of New Jersey, New Brunswick, New Jersey, USA  
^3^University of Massachusetts Dartmouth, School for Marine Science
and Technology, Massachusetts, USA  
^4^The Nature Conservancy, Rhode Island, USA  
^5^University of Rhode Island, Graduate School of Oceanography, Narragansett, RI, USA  
^6^New England Fishery Management Council, Newburyport, MA, USA  
^7^Mid-Atlantic Fishery Management Council, Dover, DE, USA  

*Add Figure 1.2 from report*  (addresses reviewer comment about distribution of Mohn's rho)
Distribution of 50 Mohn’s rho values for no misspecification (top row), catch misspecification (middle row), and natural mortality misspecification (bottom row) for spawning stock biomass (SSB), fishing mortality rate (Fbar), and recruitment (R). Panels are defined by retrospective type (none, catch, natural mortality), fishing intensity (1=always overfishing, 2=overfishing reducing to Fmsy), and number of selectivity blocks (1 or 2).

*Add Figure 1.5 from report* (addresses reviewer comment about scenario set ups)
Fishing mortality rate (F) during the base period (black line) for the 8 combinations of F history (F = Fmsy in second half of base period, O = overfishing throughout), number of fishery selectivity blocks (1 or 2), and retrospective source (catch or natural mortality, M). The horizontal blue line denotes the F at maximum sustainable yield (Fmsy) computed from conditions in the first year of the base period (1970) while the red dashed horizontal line denotes the Fmsy computed from conditions in the terminal year in the base period (2019).

*Add Figure 1.6 from report* (addresses reviewer comment about scenario set ups)
Spawning stock biomass (SSB) relative to the SSB at maximum sustainable yield (SSBmsy) computed from the conditions in the terminal year in the base period (2019) with the median value shown as the black line, the interquartile range shown in light grey, and the 95% range shown in dark grey for the 8 combinations of F history (F = Fmsy in second half of base period, O = overfishing throughout), number of fishery selectivity blocks (1 or 2), and retrospective source (catch or natural mortality, M).

*Add Figure 4.15 from report* (addresses reviewer comment about no retro results)
Mean spawning stock biomass (SSB) relative to the SSB at maximum sustainable yield (SSBmsy) and catch relative to MSY for 3 sources of retrospective pattern (catch, natural mortality (M), and none) and two fishing histories (F and A) in the long term based on 1,000 simulations.

```{r pcaplot, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat","pca.png"))
```

Figure S5. Principal components analysis of the median values of the metrics shown in Figure 1.
\newpage


```{r loop, echo=FALSE}
myfile <- list()
mytext <- list()
myfile2 <- list()
mytext2 <- list()
mycount <- 0
for (i in 1:2){
  s1 <- ifelse(i==1,"C","M")
  l1 <- ifelse(i==1,"catch","natural mortality")
  for (j in 1:2){
    s2 <- ifelse(j==1,"F","O")
    l2 <- ifelse(j==1,"reduced to","remaining at 2.5 times")
    for (k in 1:2){
      s3 <- k
      l3 <- ifelse(k==1,"was a single selectivity block","were two selectivity blocks")
      for (m in 1:2){
        s4 <- ifelse(m==1,"A","R")
        l4a <- ifelse(m==1,"no","a")
        l4b <- ifelse(m==1,"1","0.75")
        
        mycount <- mycount + 1
        myfile[[mycount]] <- paste0(s1,s2,s3,s4,"_C_vs_SSB.png")
        myfile2[[mycount]] <- paste0(s1,s2,s3,s4,"_F_vs_SSB.png")
        mytext[[mycount]] <- paste0("\nFigure S", mycount+5, ". Relationship between long-term average catch and spawning stock biomass relative to their reference points by method. Each point represents the average for years 21-40 in the feedback period for a single iteration of a scenario. The scenario shown is where ", l1, " was the source of the retrospective pattern with $F$ ", l2, " $F_{MSY}$ in the second half of the base period, there ", l3, ", and where ", l4a, " buffer was applied to the catch advice (catch multiplier = ", l4b, ").")
        mytext2[[mycount]] <- paste0("\nFigure S", mycount+5+16,". Relationship between long-term average fishing mortality rate and spawning stock biomass relative to their reference points by method. Each point represents the average for years 21-40 in the feedback period for a single iteration of a scenario. The scenario shown is where ", l1, " was the source of the retrospective pattern with $F$ ", l2, " $F_{MSY}$ in the second half of the base period, there ", l3, ", and where ", l4a, " buffer was applied to the catch advice (catch multiplier = ", l4b, ").")
      }
    }
  }
}
```

```{r p1, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[1]]))
```
`r mytext[[1]]`
\newpage

```{r p2, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[2]]))
```
`r mytext[[2]]`
\newpage

```{r p3, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[3]]))
```
`r mytext[[3]]`
\newpage

```{r p4, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[4]]))
```
`r mytext[[4]]`
\newpage

```{r p5, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[5]]))
```
`r mytext[[5]]`
\newpage

```{r p6, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[6]]))
```
`r mytext[[6]]`
\newpage

```{r p7, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[7]]))
```
`r mytext[[7]]`
\newpage

```{r p8, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[8]]))
```
`r mytext[[8]]`
\newpage

```{r p9, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[9]]))
```
`r mytext[[9]]`
\newpage

```{r p10, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[10]]))
```
`r mytext[[10]]`
\newpage

```{r p11, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[11]]))
```
`r mytext[[11]]`
\newpage

```{r p12, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[12]]))
```
`r mytext[[12]]`
\newpage

```{r p13, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[13]]))
```
`r mytext[[13]]`
\newpage

```{r p14, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[14]]))
```
`r mytext[[14]]`
\newpage

```{r p15, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[15]]))
```
`r mytext[[15]]`
\newpage

```{r p16, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[16]]))
```
`r mytext[[16]]`
\newpage

```{r p21, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[1]]))
```
`r mytext2[[1]]`
\newpage

```{r p22, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[2]]))
```
`r mytext2[[2]]`
\newpage

```{r p23, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[3]]))
```
`r mytext2[[3]]`
\newpage

```{r p24, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[4]]))
```
`r mytext2[[4]]`
\newpage

```{r p25, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[5]]))
```
`r mytext2[[5]]`
\newpage

```{r p26, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[6]]))
```
`r mytext2[[6]]`
\newpage

```{r p27, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[7]]))
```
`r mytext2[[7]]`
\newpage

```{r p28, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[8]]))
```
`r mytext2[[8]]`
\newpage

```{r p29, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[9]]))
```
`r mytext2[[9]]`
\newpage

```{r p210, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[10]]))
```
`r mytext2[[10]]`
\newpage

```{r p211, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[11]]))
```
`r mytext2[[11]]`
\newpage

```{r p212, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[12]]))
```
`r mytext2[[12]]`
\newpage

```{r p213, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[13]]))
```
`r mytext2[[13]]`
\newpage

```{r p214, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[14]]))
```
`r mytext2[[14]]`
\newpage

```{r p215, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[15]]))
```
`r mytext2[[15]]`
\newpage

```{r p216, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile2[[16]]))
```
`r mytext2[[16]]`
\newpage

