---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

\newpage

# Supplemental Materials for Data Rich but Model Resistant: An Evaluation of Data-Limited Methods to Manage Fisheries with Failed Age-based Stock Assessments

  
Christopher M. Legault^1^, John Wiedenmann^2^ Jonathan J. Deroba^1^, Gavin Fay^3^, Timothy J. Miller^1^, Elizabeth N. Brooks^1^, Richard J. Bell^4^, Joseph A. Langan^5^, Jamie M. Cournane^6^, Andrew W. Jones^1^, Brandon Muffley^7^


^1^National Marine Fisheries Service, Northeast Fisheries Science Center, Woods Hole, MA, USA  
^2^Department of Ecology, Evolution and Natural Resources, Rutgers, The State University of New Jersey, New Brunswick, New Jersey, USA  
^3^University of Massachusetts Dartmouth, School for Marine Science
and Technology, Massachusetts, USA  
^4^The Nature Conservancy, Rhode Island, USA  
^5^University of Rhode Island, Graduate School of Oceanography, Narragansett, RI, USA  
^6^New England Fishery Management Council, Newburyport, MA, USA  
^7^Mid-Atlantic Fishery Management Council, Dover, DE, USA  

some text

some more text

```{r loop, echo=FALSE}
myfile <- list()
mytext <- list()
mycount <- 0
for (i in 1:2){
  s1 <- ifelse(i==1,"C","M")
  l1 <- ifelse(i==1,"catch","natural mortality")
  for (j in 1:2){
    s2 <- ifelse(j==1,"F","O")
    l2 <- ifelse(j==1,"reduced to","remaining at 2.5 times")
    for (k in 1:2){
      s3 <- k
      l3 <- ifelse(k==1,"was a single selectivity block","were two selectivity blocks")
      for (m in 1:2){
        s4 <- ifelse(m==1,"A","R")
        l4a <- ifelse(m==1,"no","a")
        l4b <- ifelse(m==1,"1","0.75")
        
        mycount <- mycount + 1
        myfile[[mycount]] <- paste0(s1,s2,s3,s4,"_C_vs_SSB.png")
        mytext[[mycount]] <- paste0("\nRelationship between long-term average catch and spawning stock biomass relative to their reference points by method. Each point represents the average for years 21-40 in the feedback period for a single iteration of a scenario. The scenario shown is where ", l1, " was the source of the retrospective pattern with $F$ ", l2, " $F_{MSY}$ in the second half of the base period, there ", l3, ", and where ", l4a, " buffer was applied to the catch advice (catch multiplier = ", l4b, ").")
      }
    }
  }
}
```

```{r p1, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[1]]))
```
`r mytext[[1]]`
\newpage

```{r p2, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[2]]))
```
`r mytext[[2]]`
\newpage

```{r p3, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[3]]))
```
`r mytext[[3]]`
\newpage

```{r p4, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[4]]))
```
`r mytext[[4]]`
\newpage

```{r p5, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[5]]))
```
`r mytext[[5]]`
\newpage

```{r p6, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[6]]))
```
`r mytext[[6]]`
\newpage

```{r p7, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[7]]))
```
`r mytext[[7]]`
\newpage

```{r p8, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[8]]))
```
`r mytext[[8]]`
\newpage

```{r p9, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[9]]))
```
`r mytext[[9]]`
\newpage

```{r p10, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[10]]))
```
`r mytext[[10]]`
\newpage

```{r p11, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[11]]))
```
`r mytext[[11]]`
\newpage

```{r p12, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[12]]))
```
`r mytext[[12]]`
\newpage

```{r p13, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[13]]))
```
`r mytext[[13]]`
\newpage

```{r p14, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[14]]))
```
`r mytext[[14]]`
\newpage

```{r p15, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[15]]))
```
`r mytext[[15]]`
\newpage

```{r p16, echo=FALSE, message=FALSE, warnings=FALSE, out.width='7in'}
knitr::include_graphics(file.path(getwd(),"supplmat",myfile[[16]]))
```
`r mytext[[16]]`
\newpage

