---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#devtools::install_github("timjmiller/wham", dependencies=TRUE)
library(wham)
library(tidyverse)
library(furrr)
library(PlanBsmooth)
library(RcppRoll)

rscripts <- list.files(pattern = "\\.R$")
rscripts <- str_subset(rscripts, "Run_MSE", negate = TRUE)
map(rscripts, source)
#source("performance_metrics.R")
#source("wham_mse_functions.R")
#source("wham_mse_functions.R")
```

## get input object

```{r base_input}
base_input <- get_base_input()
base_input
base_input$IBM <- M_CC
base_input$adv.yr <- 2
```

## scenario specifications

```{r}
f_vecs <- read_csv("base_f_vecs.csv") %>% 
  rowid_to_column(var = "t") %>% 
  pivot_longer(names_to = "f_pattern", values_to = "f",
               cols = -t) %>% 
  separate(f_pattern, into = c("dummy", "f_pattern"),
           convert = TRUE) %>% 
  select(f_pattern, f) %>% 
  group_by(f_pattern) %>% 
  nest() %>% 
  rename("F" = data) %>% 
  I()
f_vecs

base_specs <- read_csv("base-scenarios.csv") %>% 
  left_join(f_vecs) %>% 
  mutate(F = map(F, as.matrix),
         mean_rec_pars = rep(list(base_input$mean_rec_pars),nrow(.)),
         mean_rec_pars = map2(mean_rec_pars, steepness, ~put_h(.x, .y))) %>% 
  select(-f_pattern, -steepness) %>% 
  I()
projection_specs <- read_csv("proj-scenarios.csv")
base_specs <- base_specs %>% 
  mutate(projection_specs = map(seq_len(nrow(.)),~I(projection_specs))) %>% 
  unnest(cols = c(projection_specs)) %>% 
  I()
base_specs
```

# get settings for scenarios want to run
```{r}
#run default and one change in this run
base_scens <- c(0,1)

mse_setup <- base_specs %>% 
  filter(base_scen %in% base_scens) %>% 
  select(-retro_type) %>% 
  group_by(base_scen, proj_scen) %>% 
  nest() %>% 
  rename("specs" = data) %>% 
  left_join(select(base_specs, base_scen, retro_type) %>% group_by(base_scen) %>% slice(1)) %>% 
  I()
mse_setup
```


## modify base input for chosen scenarios
```{r}
mse_setup <- mse_setup %>% 
  ungroup() %>% 
  mutate(input = rep(list(base_input), nrow(.)),
         input = map2(specs, input, ~change_input(.y, .x))) %>% 
  I()
mse_setup
```


## run the MSE

### add rows for each realization for each scenario
could possibly hardwire the seeds somehow

```{r}
#dimension problem
nbase <- nrow(mse_setup) #number of scenarios
nsim <- 10
nprojyrs <- 40

#generate seeds
sim_seeds <- tibble(isim = 1:nsim,
                  seed = sample(1:1000000,
                                nsim,
                                replace = FALSE))

#set up output tibble for each simulation
mse_sim_setup <- mse_setup %>% 
  mutate(sim_seeds = map(seq_len(nrow(.)),~I(sim_seeds))) %>% 
  unnest(cols = sim_seeds) %>% 
  I()
mse_sim_setup
```  


### run the MSE over each row of the mse_sim_setup

```{r}
#do the MSE for all simulations and scenarios
future::plan(future::multisession)
system.time(mse_output <- mse_sim_setup %>% 
  mutate(wham = purrr::pmap(list(seed, input),
                             ~do_wham_mse_sim(seed = .x,
                                              input = .y,
                                              nprojyrs = nprojyrs))))
  mutate(wham = furrr::future_pmap(list(seed, input),
                            ~do_wham_mse_sim(seed = .x,
                                             input = .y,
                                             nprojyrs = nprojyrs))))

#saveRDS(mse_output, file = "demo-mse-20200723.rds")
#saveRDS(mse_output, file = "2020-08-20_40sims-mse_output.rds")
```

pull out some parts of the output
calculate some performance metrics

```{r}
mse_results <- mse_output %>% 
  mutate(om_ssb = map(wham,
                  ~pluck(.x$sim_data_series$SSB)),
         catch = map(wham, 
                  ~pluck(.x$sim_data_series$catch)),
         catch = map(catch, na_if, y = "NaN"),
         om_ssb = map(om_ssb, na_if, y = "NaN"),
         refpts = map(wham, "refpts"),
         ssb_metrics = pmap(list(om_ssb, refpts), get_ssb_metrics, nprojyrs = nprojyrs),
         catch_metrics = pmap(list(catch, refpts), get_catch_metrics, nprojyrs = nprojyrs))

saveRDS(mse_results, file = "2020-08-20_40sims_mse-results.rds")
```

`mse_results` is a tibble containing the results, and vectors of the OM SSB, and a list of SSB performance metrics (fudging using SSBlim as SSBMSY for this demo)

pull out the ssb metrics

```{r}
ssb_results <- mse_results %>% 
  select(base_scen, proj_scen, isim, ssb_metrics) %>% 
  #select(iscen, isim, ssb_metrics) %>% 
  mutate(ssb_metrics = map(ssb_metrics, enframe)) %>% 
  unnest() %>% 
  mutate(value = map_dbl(value, I)) %>% 
  rename(metric = name) %>% 
  I()
ssb_results
```

summarize across simulations by scenario
25%, 50%, 75% quantiles
```{r}
quibble <- function(x, q = c(0.25, 0.5, 0.75)) {
  tibble(x = quantile(x, q, na.rm = TRUE), q = q)
}

ssb_summary <- ssb_results %>% 
  group_by(metric, base_scen, proj_scen) %>% 
  summarise(y = list(quibble(value, c(0.25, 0.5, 0.75)))) %>% 
  tidyr::unnest(y) %>% 
  I()
ssb_summary
```

